Safe Memory Regions for Big Data Processing

### 2016-05-31 paper/type-inference.tex
Updates to type inference description

@@

% Validity constraints primarily result from checking if a region type
% template of form $B\inang{\tbar}\inang{\rhoalloc,\rhobar}$ is
% well-formed (i.e., $\typeOk$). The corresponding well-formedness rule
% (Fig.~\ref{fig:fb-staticsem}) requires that the instantiation of $B$'s
% formal region parameters with actual region arguments satisfy its
% precondition, where actuals are subsituted for formals. Since $B$'s
% precondition could be unknown, the requirement results in a in a
% validity constraint with predicate variables and pending
% substitutions. Other sources of validity constraints are calls to
% region-polymorphic methods, where similar situation arises. 
In general, validity constraints generated by our algorithm assume one
of the following two forms:
\begin{center}
\( \isvalid{\phictxt \conj \bigwedge_{i}\varphi_i}
           {\phicstr} \qquad
   \isvalid{\phictxt \conj \bigwedge_{i}\varphi_i}
           {F_j(\varphi_j)}
\)
\end{center}
Where $\varphi_i$'s denote the unknown preconditions of the class and
the method under elaboration. If the constraint is generated while
checking the well-formedness of a type or elaborating an expression,
then $\varphi_j$'s denote the unknown preconditions of classes and
methods that were used in that type or expression.  Each use of a
(region-polymorphic) class or a method may instantiate region
parameters differently, resulting in a different pending substitution
($F_j$). Formulas $\phictxt$ and $\phicstr$ are concrete, i.e., free
of predicate variables and pending substitutions. While $\phictxt$
captures relationships that are \emph{known} to hold between concrete
region identifiers (i.e., $\rgn$'s) when the constraint was generated,
$\phicstr$ captures relationships that are \emph{required} to hold
among region varibles (i.e., $\rho$'s), or relationships between
region variables and identifiers. Each region variable occuring in
$\phicstr$ has an associated well-formedness constraint, which
specifies its unification domain. The unification domain of a
constraint is the union of unification domains of all region variables
occuring in the constraint.

@@

Given a set of validity constraints, we first build a dependency graph
($G_c$) with constraints as nodes, and dependencies between them
captured as edges. There exists an edge from a constraint $c_2$ to a
constraint $c_1$ in the graph (i.e., $c_2$ \emph{depends on} $c_1$) if
any of the following conditions hold:
\begin{itemize}
\item $c_1$ constrains a predicate variable that $c_2$ uses. 
\item $c_1$ constrains a (non-strict) subset of predicate variables
that $c_2$ constraint.
\end{itemize}
The first condition intuitively corresponds to a case, where
expression or type, whose region elaboration is constrained by $c_2$
refers to a method or a class, whose unknown precondition is
constrained by $c_1$. The common predicate variable ($\varphi$)
represents the unknown precondition in this case. The dependency from
$c_2$ to $c_1$ means that $c_1$ must be solved to compute $\varphi$
before $c_2$ is solved, thus enforcing the rule that the precondition
of a method must not depend on its calling context. 
% Note that the first condition results in self-loops over recursive
% constraints in the dependency graph.  Since mutually recursive
% definitions refer each other, the first condition also results in
% bidirectional dependencies between mutually recursive constraints
% (and self-loops over recursive constraints). 
The second condition captures two kinds of dependencies. First, the
dependency from the constraints over a method precondition to the
constraints over the precondition of the class containing the method.
This captures our preference that the constraints over a class's
region parameters should not depend on the idiosyncracies of its
methods. Any additional constraints required by any of its methods
must be captured in the precondition of the method itself
(well-formedness rules allow this possibility). The second condition
adds bidirectional dependencies between validity constraints that
constraint the same set of predicate variables.

@@

\paragraph{Example} The dependency DAG ($G_C$) over validity
constraints from the \C{Pair} example (\S~\ref{sec:constraints}) is
shown below: %in Fig.~\ref{fig:pair-deps}. 
\vspace*{-0.08in}
\begin{figure}[H]
\includegraphics[scale=0.6]{DepDAG.png}
\end{figure}
\vspace*{-0.08in}
Each node (labeled $C_i$) is a set of constraints that belong to a
strongly connected component in the dependency graph ($G_c$), hence
are mutually dependent. All dependencies, except the self-dependency
on $\mathbf{c_{10}}$, are type-2 dependencies.

@@

For the DAG in figure above, we consider the topological order $[C_1,
C_2, C_3]$ of its transpose, and solve the sets of constraints in that
order.  Solving the set ($C$) of constraints entails finding an
assignment for $C$'s subject ($\varphi$), and also all the region
variables ($\rhobar$) that occur free in $C$, such that the solution
satisfies well-formedness constraints on $\varphi$ and $\rhobar$. To
simplify presentation, we think of $C$ as being parameterized on
$\varphi$ and $\rhobar$, and write it as $C\lbrack \varphi,\rhobar
\rbrack$. We now formalize the constraint satisfaction problem, and
its solution.

>>>>

\input{fb-constraint-generation}

@@

% Validity constraints primarily result from checking if a region type
% template of form $B\inang{\tbar}\inang{\rhoalloc,\rhobar}$ is
% well-formed (i.e., $\typeOk$). The corresponding well-formedness rule
% (Fig.~\ref{fig:fb-staticsem}) requires that the instantiation of $B$'s
% formal region parameters with actual region arguments satisfy its
% precondition, where actuals are subsituted for formals. Since $B$'s
% precondition could be unknown, the requirement results in a in a
% validity constraint with predicate variables and pending
% substitutions. Other sources of validity constraints are calls to
% region-polymorphic methods, where similar situation arises. 
In general, validity constraints generated by our algorithm assume one
of the following two forms:
\begin{center}
\( \isvalid{\phictxt \conj \varphi_i}
           {\phicstr} \qquad
   \isvalid{\phictxt \conj \varphi_i}
           {F_j(\varphi_j)}
\)
\end{center}
where $\varphi_i$ and $\varphi_j$ denote the unknown preconditions of some
methods or constructors.
%
%% If the constraint is generated while checking the well-formedness of a
%% type or elaborating an expression, then $\varphi_j$'s denote the unknown
%% preconditions of classes and methods that were used in that type or expression.
%
Each use of a (region-polymorphic) class or a method may instantiate region
parameters differently, resulting in a different pending substitution
($F_j$). Formulas $\phictxt$ and $\phicstr$ are concrete, i.e., free
of predicate variables and pending substitutions. While $\phictxt$
captures relationships that are \emph{known} to hold between concrete
region identifiers (i.e., $\rgn$'s) when the constraint was generated,
$\phicstr$ captures relationships that are \emph{required} to hold
among region varibles (i.e., $\rho$'s), or relationships between
region variables and identifiers.
%
Each region variable occuring in $\phicstr$ has an associated well-formedness
constraint, which specifies its unification domain. The unification domain of a
constraint is the union of unification domains of all region variables
occuring in the constraint.

@@

Given a set of validity constraints, we first build a dependency graph
($G_c$) with constraints as nodes, and dependencies between them
captured as edges. There exists an edge from a constraint $c_2$ to a
constraint $c_1$ in the graph (i.e., $c_2$ \emph{depends on} $c_1$) if
$c_1$ constrains a predicate variable that $c_2$ uses.

The above condition intuitively corresponds to a case, where
expression or type, whose region elaboration is constrained by $c_2$
refers to a method or a class, whose unknown precondition is
constrained by $c_1$. The common predicate variable ($\varphi$)
represents the unknown precondition in this case. The dependency from
$c_2$ to $c_1$ means that $c_1$ must be solved to compute $\varphi$
before $c_2$ is solved, thus enforcing the rule that the precondition
of a method must not depend on its calling context. 
% Note that the first condition results in self-loops over recursive
% constraints in the dependency graph.  Since mutually recursive
% definitions refer each other, the first condition also results in
% bidirectional dependencies between mutually recursive constraints
% (and self-loops over recursive constraints).

@@

%% \paragraph{Example} The dependency DAG ($G_C$) over validity
%% constraints from the \C{Pair} example (\S~\ref{sec:constraints}) is
%% shown below: %in Fig.~\ref{fig:pair-deps}. 
%% \vspace*{-0.08in}
%% \begin{figure}[H]
%% \includegraphics[scale=0.6]{DepDAG.png}
%% \end{figure}
%% \vspace*{-0.08in}
%% Each node (labeled $C_i$) is a set of constraints that belong to a
%% strongly connected component in the dependency graph ($G_c$), hence
%% are mutually dependent. All dependencies, except the self-dependency
%% on $\mathbf{c_{10}}$, are type-2 dependencies.

@@

%% For the DAG in figure above, we consider the topological order $[C_1,
%% C_2, C_3]$ of its transpose, and solve the sets of constraints in that
%% order.

Solving the set ($C$) of constraints entails finding an
assignment for $C$'s subject ($\varphi$), and also all the region
variables ($\rhobar$) that occur free in $C$, such that the solution
satisfies well-formedness constraints on $\varphi$ and $\rhobar$. To
simplify presentation, we think of $C$ as being parameterized on
$\varphi$ and $\rhobar$, and write it as $C\lbrack \varphi,\rhobar
\rbrack$. We now formalize the constraint satisfaction problem, and
its solution.
